---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
df <- read.csv("activity.csv", colClasses=c("integer", "character", "integer"))
df$date <- as.Date(df$date, format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r}
total <- with(df, aggregate(steps~date, FUN=sum, na.rm=TRUE))
hist(total$steps)
meanSteps <- as.integer(mean(total$steps, na.rm=TRUE))
medianSteps <- median(total$steps, na.rm=TRUE)
```
The mean of steps taken per day is `r meanSteps` and the median of steps taken per day is `r medianSteps`

## What is the average daily activity pattern?
```{r}
timeSeries <- with(df, aggregate(steps ~ interval, FUN=mean, na.rm=TRUE))
with(timeSeries, plot(interval, steps, type="l", main="Average Daily Activity Pattern"))
stepsMax <- as.integer(max(timeSeries$steps))
intervalMax <- timeSeries[which.max(timeSeries$steps),1]
```
Averaged across all the days in the dataset, the maximum number of steps is `r stepsMax` and is on interval `r intervalMax`.


## Imputing missing values
```{r}
totalNA <- sum(is.na(df$steps))
meanFiller <- as.integer(mean(df$steps, na.rm=TRUE))
newDF <- data.frame(df)
newDF[is.na(df$steps),1] <- meanFiller
sum(is.na(newDF))
newTotal <- with(newDF, aggregate(steps~date, FUN=sum, na.rm=TRUE))
hist(newTotal$steps)
meanSteps <- mean(newTotal$steps, na.rm=TRUE)
medianSteps <- median(newTotal$steps, na.rm=TRUE)
```
The missing data were replaced with the average steps for all days and interval. 
Averaged across all the days in the dataset, the maximum number of steps is `r stepsMax` and is on interval `r intervalMax`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
newDF$week <- weekdays(newDF$date)
newDF$week[!(newDF$week == 'Saturday'|newDF$week == 'Sunday')] <- 'weekday'
newDF$week[(newDF$week == 'Saturday'|newDF$week == 'Sunday')] <- 'weekend'
library(plyr)
library(lattice)
week <- ddply(newDF, c("interval", "week"), function(x) apply(x[1],2, mean))
head(week)
xyplot(steps ~ interval | week, data = week, type="l", xlab="Interval", ylab="# of Steps", layout = c(1,2))
```




